Gravitational Lens Recovery with \Glass: Measuring the mass profile and shape of a lens

### 2014-01-22 limits/ms.tex
eqnref -> eqnrefp in appropriate places.

\subsection{Discrete models}\label{sec:discrete}
For this paper we will restrict ourselves to using the pixelated basis set that
has been employed in previous work and is implemented in \PixeLens. Here we
will briefly review the main ideas. The algorithm for generating models in
\Glass\ samples a convex polytope in a high dimensional space whose interior
points are solutions to the lens equation and satisfy other physically
motivated linear priors (the sampling algorithm in \Glass\ is \emph{not} the same as
employed in \PixeLens).  We therefore formulate all of our equations as
equations linear in the unknowns. We describe the density distribution $\kappa$
as a set of discrete grid cells or pixels $\kappa_i$ and rewrite the potential
(\eqnref{lensing potential}) as
%
\begin{equation}
  \psi(\vec\theta) = \sum_n \kappa_n Q_n(\vec\theta)
  \label{discrete potential}
\end{equation}
%
where the sum runs over all the pixels and $Q_n$ is the integral of the logarithm
over pixel $n$. The exact form for $Q$ is described in \appref{Q derivation}.
We can find the discretized lens equation by simply taking the gradient of the
above equations.

@@

\subsection{Raytracing}\label{Raytracing}
\Glass\ can also determine the position of images and time delays from 
particle-based simulation output given a source position $\vec\beta$. This is
used to generate the lens configurations used in the parameter study.  The
particles are first projected onto a very high resolution grid representing the
lens plane. The centers $\vec\theta_i$ of each of the grid cells are mapped
back onto the source plane using \eqnref{eqn:lens_equation}. If the location on
the source plane $\vec\beta_i$ is within a user specified
$\eps_\mathrm{accept}$ of $\vec\beta$ then $\vec\theta_i$ is 
accepted and further refined using a root finding algorithm until the distance
to $\vec\beta$ is nearly zero. If multiple points converge to an
$\eps_\mathrm{root}$ of each other then only one point is taken.  Care must be
taken that the grid resolution is high enough that the resulting image position
error is below the equivalent observational error. Time delays are then
calculated in order of the arrival time at each image (\eqnref{tau}).

@@

The stellar mass distribution gives a lower bound on the total mass. Where the stars dominate the central potential, it can provide a powerful constraint extra to the strong lensing data. We took the stellar mass directly from
the generated galaxies and projected the particles onto the pixels. \Glass\
also offers an option to interpolate any map of stellar mass (e.g., from an
observation) onto the pixels. The linear constraint is added to \Glass\ by
writing $\kappa_n = \kappa_{dm,n} + \kappa_{s,n}$ as the sum of the
dark matter and stellar mass components in the potential (\eqnref{discrete
potential}). Since each $\kappa_{s,n}$ is just a constant we do not add new,
separate equations for each pixel. Although we assume a perfect recovery of the stellar mass here, it is straightforward to add errors as the stellar mass constraint remains linear: $\kappa_n = \kappa_{dm,n} + \epsilon \kappa_{s,n}$, where $\epsilon \sim 1$ is an additional error parameter.

@@

As outlined in \S\ref{sec:glass}, \Glass\ can also run post processing routines
on the model ensemble which can be used to apply non-linear constraints. As an
example, we consider here constraints from stellar kinematics. The models in
the \Glass\ ensemble are processed as described in \S\ref{sec:glasskinematics}.
To illustrate the power of stellar kinematic constraints, in
\figref{sigma-beta}, we plot the projected velocity dispersion calculated for
one model model (extracted from the full ensemble) of the BC Quad with time
delays and stellar mass (left), and the mean projected velocity dispersion
(\eqnref{eqn:virialaverage}) as a function of the upper limit of the averaging
integral (the aperture size), $a_p$ (right). In both cases, we calculate curves
for two extrema velocity anisotropies: $\beta=0$ (green) and $\beta=1$ (red).
Over-plotted is the correct answer for the BC model (black). The stellar half
mass radius (yellow) Einstein radius (black) are marked by vertical lines. For
this configuration, these two radii are well-separated.

>>>>

\subsection{Discrete models}\label{sec:discrete}
For this paper we will restrict ourselves to using the pixelated basis set that
has been employed in previous work and is implemented in \PixeLens. Here we
will briefly review the main ideas. The algorithm for generating models in
\Glass\ samples a convex polytope in a high dimensional space whose interior
points are solutions to the lens equation and satisfy other physically
motivated linear priors (the sampling algorithm in \Glass\ is \emph{not} the same as
employed in \PixeLens).  We therefore formulate all of our equations as
equations linear in the unknowns. We describe the density distribution $\kappa$
as a set of discrete grid cells or pixels $\kappa_i$ and rewrite the potential
\eqnrefp{lensing potential} as
%
\begin{equation}
  \psi(\vec\theta) = \sum_n \kappa_n Q_n(\vec\theta)
  \label{discrete potential}
\end{equation}
%
where the sum runs over all the pixels and $Q_n$ is the integral of the logarithm
over pixel $n$. The exact form for $Q$ is described in \appref{Q derivation}.
We can find the discretized lens equation by simply taking the gradient of the
above equations.

@@

\subsection{Raytracing}\label{Raytracing}
\Glass\ can also determine the position of images and time delays from 
particle-based simulation output given a source position $\vec\beta$. This is
used to generate the lens configurations used in the parameter study.  The
particles are first projected onto a very high resolution grid representing the
lens plane. The centers $\vec\theta_i$ of each of the grid cells are mapped
back onto the source plane using \eqnref{eqn:lens_equation}. If the location on
the source plane $\vec\beta_i$ is within a user specified
$\eps_\mathrm{accept}$ of $\vec\beta$ then $\vec\theta_i$ is 
accepted and further refined using a root finding algorithm until the distance
to $\vec\beta$ is nearly zero. If multiple points converge to an
$\eps_\mathrm{root}$ of each other then only one point is taken.  Care must be
taken that the grid resolution is high enough that the resulting image position
error is below the equivalent observational error. Time delays are then
calculated in order of the arrival time at each image \eqnrefp{tau}.

@@

The stellar mass distribution gives a lower bound on the total mass. Where the stars dominate the central potential, it can provide a powerful constraint extra to the strong lensing data. We took the stellar mass directly from
the generated galaxies and projected the particles onto the pixels. \Glass\
also offers an option to interpolate any map of stellar mass (e.g., from an
observation) onto the pixels. The linear constraint is added to \Glass\ by
writing $\kappa_n = \kappa_{dm,n} + \kappa_{s,n}$ as the sum of the
dark matter and stellar mass components in the potential \eqnrefp{discrete
potential}. Since each $\kappa_{s,n}$ is just a constant we do not add new,
separate equations for each pixel. Although we assume a perfect recovery of the stellar mass here, it is straightforward to add errors as the stellar mass constraint remains linear: $\kappa_n = \kappa_{dm,n} + \epsilon \kappa_{s,n}$, where $\epsilon \sim 1$ is an additional error parameter.

@@

As outlined in \S\ref{sec:glass}, \Glass\ can also run post processing routines
on the model ensemble which can be used to apply non-linear constraints. As an
example, we consider here constraints from stellar kinematics. The models in
the \Glass\ ensemble are processed as described in \S\ref{sec:glasskinematics}.
To illustrate the power of stellar kinematic constraints, in
\figref{sigma-beta}, we plot the projected velocity dispersion calculated for
one model model (extracted from the full ensemble) of the BC Quad with time
delays and stellar mass (left), and the mean projected velocity dispersion
\eqnrefp{eqn:virialaverage} as a function of the upper limit of the averaging
integral (the aperture size), $a_p$ (right). In both cases, we calculate curves
for two extrema velocity anisotropies: $\beta=0$ (green) and $\beta=1$ (red).
Over-plotted is the correct answer for the BC model (black). The stellar half
mass radius (yellow) Einstein radius (black) are marked by vertical lines. For
this configuration, these two radii are well-separated.
